Galaxy Clustering Constraints on the Galaxy Size--Halo Connection

### 2017-09-02 paper/galsize_paper.tex
Minor changes and typo corrections

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-mass radius $\rhalf$ of galaxy bulges and disks. We show that both $\rhalfdisk$ and $\rhalfbulge$ are well-described by power law scaling relations with halo virial radius, $\rhalf=A\rvir^{\alpha}.$ Novel to this work, we use new SDSS measurements of the $\rhalftot-$dependence of galaxy clustering to constrain the model parameters, $A_{\rm bulge}, A_{\rm disk}, \alpha_{\rm bulge},\alpha_{\rm disk},$ and log-normal scatter $\sigmarhalf.$ Even when only coarsely tuning these parameters to the observed one-point functions $\langle\rhalfdisk\vert\mstartot\rangle$ and $\langle\rhalfbulge|\mstartot\rangle,$ our model accurately predicts the observed two-point clustering on small- and large-scales. This success non-trivial, as we show that galaxy clustering is highly sensitive to the physics that shapes satellite galaxy profiles. We find no evidence for the commonly assumed relation between halo spin $\halospin$ and $\rhalfdisk,$ and show that this assumption cannot be meaningfully constrained with either the clustering or lensing of  $L_{\ast}$ galaxies. Our results provide simple boundary conditions for more complex and fine-grained models of galaxy size. We make our python code publicly available to support cosmological surveys that require realistic synthetic galaxy populations.
\end{abstract}

@@

In all sections that follow, we will declutter our notation by referring to $\rhalftot$ as $\rhalf$ and $\mstartot$ as $\mstar.$ In all cases where we wish to refer to the half-light radius or stellar mass of a particular component, that component will explicitly appear in superscript. All numerical values of $\rhalf$ will be quoted in $\kpc,$ and all values of $\mstar$ and $\mhalo$ in $\msun,$ assuming $H_0=67.8~\kms\equiv100h~\kms$ the best-fit value from \citet{planck15}. To scale stellar masses to ``$h=1$ units" \citep{croton13}, our numerically quoted values for $\mstar$ should be multiplied by a factor of $h^2,$ while our halo masses and distances should be multiplied by a factor of $h.$

@@

The starting point of our model is the best-fit {\tt UniverseMachine} model \citep[][in prep]{behroozi_etal17}. This model maps stellar mass and star-formation rate to every halo and subhalo at each snapshot in the Bolshoi-Planck simulation. The stellar mass function, quenched fraction, SFR density, and SFR-dependent clustering are all accurately captured by this model from $z=0-10.$ The model we have developed in the present work does not depend in an essential way upon the {\tt UniverseMachine} model in particular, we only require some reasonably accurate starting point for the $\mstar$ and $\ssfr$ values mapped to simulated halos.

@@

For mock galaxies, to compute galaxy clustering we employ the distant observer approximation by treating the simulation $z-$axis as the line-of-sight. We compute $\wproj$ using the {\tt mock\_observables.wp} function in {\tt Halotools}, which is a python implementation of the algorithm in the {\tt Corrfunc} C library \citep{sinha_etal17}.

>>>>

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-mass radius $\rhalf$ of galaxy bulges and disks. We show that both $\rhalfdisk$ and $\rhalfbulge$ are well-described by power law scaling relations with halo virial radius, $\rhalf=A\rvir^{\alpha}.$ Novel to this work, we use new SDSS measurements of the $\rhalf-$dependence of galaxy clustering to constrain the model parameters, $A_{\rm bulge}, A_{\rm disk}, \alpha_{\rm bulge},\alpha_{\rm disk},$ and log-normal scatter $\sigmarhalf.$ Even when only coarsely tuning these parameters to the observed one-point functions $\langle\rhalfdisk\vert\mstartot\rangle$ and $\langle\rhalfbulge|\mstartot\rangle,$ our model accurately predicts the observed two-point clustering on small- and large-scales. This success non-trivial, as we show that galaxy clustering is highly sensitive to the physics that shapes satellite galaxy profiles. We find no evidence for the commonly assumed relation between halo spin $\halospin$ and $\rhalfdisk,$ and show that this assumption cannot be meaningfully constrained with either the clustering or lensing of  $L_{\ast}$ galaxies. Our results provide simple boundary conditions for more complex and fine-grained models of galaxy size. We make our python code publicly available to support cosmological surveys that require realistic synthetic galaxy populations.
\end{abstract}

@@

In all sections that follow, we will declutter our notation by referring to $\rhalftot$ as $\rhalf$ and $\mstartot$ as $\mstar.$ In all cases where we wish to refer to the half-light radius or stellar mass of a particular component, that component will explicitly appear in superscript. All numerical values of $\rhalf$ will be quoted in $\kpc,$ and all values of $\mstar$ and $\mhalo$ in $\msun,$ assuming $H_0=67.8~\kms\equiv100h~\kms,$ the best-fit value from \citet{planck15}. To scale stellar masses to ``$h=1$ units" \citep{croton13}, our numerically quoted values for $\mstar$ should be multiplied by a factor of $h^2,$ while our halo masses and distances should be multiplied by a factor of $h.$

@@

The starting point of our model is the best-fit {\tt UniverseMachine} model \citep[][in prep]{behroozi_etal17}. This model maps stellar mass and star-formation rate to every halo and subhalo at each snapshot in the Bolshoi-Planck simulation. The stellar mass function, quenched fraction, SFR density, and SFR-dependent clustering are all accurately captured by this model from $z=0-10.$ The model we have developed in the present work does not depend in an essential way upon the {\tt UniverseMachine} model in particular; we only require some reasonably accurate starting point for the $\mstar$ and $\ssfr$ values mapped to simulated halos.

@@

For mock galaxies, to compute galaxy clustering we employ the distant observer approximation by treating the simulation $z-$axis as the line-of-sight. We compute $\wproj$ using the {\tt mock\_observables.wp} function in {\tt Halotools}, which is a python implementation of the algorithm in the {\tt Corrfunc} C library \citep{sinha_etal17}. We compute lensing $\Delta\Sigma$ using the {\tt mock\_observables.delta\_sigma} function in {\tt Halotools} to calculate the two-dimensional surface mass density of a random downsampling of $10^6$ dark matter particles in the snapshot.
